# Lesson 1 Getting Ready {#L1}

## Get R and Rstudio

We need to

1. Download R from [CRAN](https://cran.r-project.org/)
1. Install R---just use the default settings
1. Download RStudio from [RStudio/posit](https://posit.co/downloads/)
1. Install RStudio

## Use RStuio to install R packages

Open RStudio. Click on `File>New File>R Script`, next copy the following R code and paste it to the **screen** called `Untitled1`

```{r, eval=FALSE}
install.packages(c("dplyr", "tidyr"))
install.packages("ggplot2")
install.packages("data.table")
```

Run each line of code by i) moving the cursor in that line; ii) pressing down **Ctrl** key then clicking on **Enter** key

## Explore Rstudio

**Tasks** :

- Figure out how to create a new project, called `my_proj_1`

- Figure out how to create an R script, called `try_1.R`, which is under the just created project `my_proj_1` and  contains the following R code
```{r, eval=FALSE}
1+1
print("Hello World!")
rnorm(5)
```

- Figure out how to run R code in `try_1.R`: a) line by line; b) all three lines "together".


