# Lesson 3 R Essentials {#L3}

## A formula

Learning any a computer language, we aim to write programs for solving problems. It's useful to know the following formula:
$$
\hbox{program}\approx \hbox{algorithm} + \hbox{data structure} + \hbox{language}
$$

**Algorithm**, roughly speaking, is about **logic steps** towards solving a problem, but it is beyond the scope of this book. Fortunately, at a beginning level, all the algorithms that we need are pretty easy.

We will cover main **data structures** in R in Lesson \@ref(L4).

As for **language**, specifically R language, we will grasp the most important things in this lesson.

## R language

To learn R, we must learn

1. assignment
1. `if else`
1. `for` loop
1. built-in R functions
1. writing R functions
1. R packages

Note that items 1 to 5 in the above are in the domain of **base R**; various R packages that we install after having R can be referred to as **extended R**. In this section, we will cover items 1 to 3; items 4 and 5 will be the topics for Lesson \@ref(L5); item 6 will be covered in Lesson \@ref(L6).

### Assignment statements 

In R, an assignment statement creates a new (vector) variable and gives it value(s).

**Examples:**
```{r}
my_name <- "Larry"
age <- 25
sex <- "male"
the_nbrs <- 1:100
five_random_nbrs <- c(1.2, 2.8, 5.9, 4.0, 3.7)
a_flag <- c(TRUE, FALSE, TRUE)
```

As shown in the above examples, an assignment is something like
```{r, eval=FALSE}
variable_name <- values
```

**Rules:**

1. A variable name contains letters, digits (0-9), `.` (dot) or `_` (underscore).
1. A variable name normally starts with a letter.
1. Reserved words (e.g. `if`) cannot be used as a variable name.

### `if else`

**Examples:**

```{r}
x <- 10
if(x > 5) print("x is greater than 5.")
if(x > 50) {print("x is greater than 50")
} else {print("x is NOT greater than 50")}
```

**Rules:**

- `if(conditional) {execution}`
- `if(conditional) {execution} else {alternative execution}`


**`else if`**: this is kind of special. Let's have an example. We create a function for giving a score, either 'green', 'orange', or 'red', based on `x` is positive and $\le$ 1, greater than 1 but $\le$ 2, or greater than 2. 

```{r}
have_a_score <- function(x) 
{if(x < 0) {
   "invalid"
} else if(x <= 1) {
   "green"
} else if(x <= 2) {
   "orange"
} else {
   "red"
}
}

print(have_a_score(-1))
print(have_a_score(0.5))
```

**`ifelse`**: it is a vectorized version of `if else`. Example:
```{r}
x <- 1:10
output <- ifelse(x %% 2 == 0, "even", "odd")
print(output)
```



### `for` loop

**Examples:** 

```{r}
s <- 0
for(i in 1:100) {
  s <- s + i
}
print(s)

for(a in letters[1:5]) {
  print(a)
}

```

**Rules:**

- `for(a_var in a_vector) {action}`

