# Lesson 2 Toy Examples {#L2}

There are three toy examples; go through them one by one.
Firstly, think about how to solve the problem by yourself; and then study the example R code.  

**Example 1:** Write an R program to plot $f(x)=\sin(x)$ (black color) and $g(x)=\cos(x)$ (red color) together; $x$ takes values from $-2\pi$ to $2\pi$.

```{r}
x <- seq(-2*pi, 2*pi, by = 0.01)
y_1 <- sin(x)
y_2 <- cos(x)
plot(x, y_1, type = 'l', col = 'black', xlab = 'x', ylab = 'y',
     main = "Plot of sin(x) (black) and cos(x) (red)")
points(x, y_2, type = 'l', col = 'red')
abline(h = 0, lty = 2)
abline(v = 0, lty = 2)
```

**Example 2:** Check all integers from 1 to 100: if an integer is multiple of 5, then output "F"; if the integer is multiple of 7, then output "S"; if it's multiple of both 5 and 7, then output "FS"; otherwise just output the integer.

```{r}
nbr <- 1:100
output <- rep("", times = 100)
for(i in nbr) {
  s <- ""
  if(i %% 5 == 0) s <- "F"
  if(i %% 7 == 0) s <- paste0(s, "S")
  if(s != "") {output[i] <- s
  } else {output[i] <- i}
}
print(output)
```

**Example 3:** Write an **R function** to calculate how old you are in days.

```{r, message=FALSE, warning=FALSE}
library(lubridate) # to use function days()
how_old_in_days <- function(`birth_day_yyyy-mm-dd`)
{today <- Sys.Date()
 life_in_days <- days(today - as.Date(`birth_day_yyyy-mm-dd`))
 print(life_in_days)
}
how_old_in_days("2005-12-04") # today is "2022-12-25"
```

