# Lesson 5 R Functions {#L5}

Keep this in mind: In R, we use functions to get work done. So learning R in some a sense is learning how to use existed R functions and how to create new R functions. Existed R functions are those included in base R and extended R packages. In this lesson, we firstly list some built-in-base-R functions, and then touch upon how to write our own R functions.

## Built-in Base R Functions

Function       |Description           | Example
:--------------|:---------------------|:-------------
`:`            |create integer vector | `x <- 1:5`
`c()`          |for creating vector   | `x <- c(1, 5, 8)`
`rep()`        |for creating vector   | `x <- rep(0, 100)`
`seq()`        |for creating vector   | `x <- seq(0, 1, by = 0.2)`
`length()`     |get vector's length   | `n <- length(x)`
`print()`      |print out             | `print(x)`
`list()`       |create a list         | `my_list <- list(a = 1, b = 2:6)`
`matrix()`     |create a matrix       | `m <- matrix(1:9, 3, 3)`
`typeof()`     |get vector's type     | `typeof(x)`
`plot()`       |for plotting          |
`getwd()`      |get working directory | `getwd()`
math functions |                      | `abs(x)`; `exp(x)`
stats functions|                      | `mean(x)`; `sd(x)`;`rnorm(10)`

The following code allows us to see many more built-in-base-R functions:
```{r, eval=FALSE}
# Source: https://stackoverflow.com/questions/58476696/list-of-all-functions-in-base-r
base_packages  <-  getOption('defaultPackages')
names(base_packages)  <-  base_packages
lapply(base_packages, function (pkg) ls(paste0('package:', pkg)))
```


## Wrting R Functions

### Basics 

An R function looks like
```{r, eval=FALSE}
func_name <- function(arguments)
{
  body
}
```

Some points:

- "Generally, function names should be verbs, and arguments should be nouns." (Wickham and Grolemund, https://r4ds.had.co.nz/functions.html) This is a good convention; anyway, function names should make sense and often should be self explained.
- The body part can have many lines of R code; the last line is often `return(an_object)` (`an_object` can be an atomic vector, matrix, data frame or list) 

### Tips 

- Use "divide and conquer" strategy to break a big task into small ones, and **write a function for each small task rather than write a huge function** for the big task. The advantages of this approach are: a) easier for debugging and maintenance; b) some small functions may be useful for other tasks.
- Use `lapply()` and other helper functions (e.g. `unlist()`) to give a vectorized function---see this in the following example, "get sum of words".
- Using `match.arg()`---see this in the following example, "add, subtract and multiply".
- Function factory---see this in the following example, "create an operation".

**Example:** get sum of words
```{r}
letter_to_nbr <- function(a_letter) 
{which(letters %in% a_letter)
}
get_sum_of_a_word <- function(a_word)
{chars <- unlist(strsplit(a_word, split = ""))
 the_nbrs <- vapply(chars, letter_to_nbr, numeric(1))
 return(sum(the_nbrs))
}
get_sum_of_words <- function(word_vec)
{unlist(lapply(word_vec, get_sum_of_a_word))
}
# testing
get_sum_of_a_word("aabbcde")
get_sum_of_words(c('a', 'ab', 'abc'))
```

**Example:** using `match.arg()`
```{r}
add_subtr_multi <- function(x, y, method = c("sum", "subtract", "multipliction"))
{method <- match.arg(method)
 switch(method,
        "sum" = x + y,
        "subtract" = x - y,
        "multipliction" = x * y)
}
# testing
add_subtr_multi(1, 2, "sum")
add_subtr_multi(2, 1, "subtract")
add_subtr_multi(10, 10, "multipliction")
```


**Example:** function factory
```{r}
create_an_op <- function(op = c("sum", "subtract", "multipliction"))
{op <- match.arg(op)
 function(x, y) {
   switch(op,
          "sum" = x + y,
          "subtract" = x - y,
          "multipliction" = x * y)
           
 }
}

# testing
plus_op <- create_an_op("sum")
subtr_op <- create_an_op("subtract")
multi_op <- create_an_op("multipliction")
plus_op(1, 2)
subtr_op(1, 2)
multi_op(10, 10)
```

