# Lesson 6 R Packages {#L6}

An R package basically is package of R functions plus metadata and documentations.

```{r, include=FALSE}
# base packages
x <- sessionInfo()[[6]][c(7, 6, 1, 4, 2, 3, 5)]
```

Base R (the R that we download from CRAN) has at least these R packages

- `r x[1]`
- `r x[2]`
- `r x[3]`
- `r x[4]`
- `r x[5]`
- `r x[6]`
- `r x[7]`

We call all the other packages that not included in base R **extended R packages**. In this lesson, we will very briefly introduce some extended R packages, and then touch upon writing R packages.

## Extended R Packages

package name      | remarks
:-----------------|:-------------------------------
`tidyverse`       | a set of packages; for data manipulation/analysis/visualization
`data.table`      | for data manipulation; enabling fast code
`lubridate`       | for dealing with dates, times etc.
`shiny`           | for creating shiny apps
`bookdown`        | for writing books
`blogdown`        | for creating web sites
`devtools`        | for creating R packages

**NB:** We can see all packages included in `tidyverse`
```{r}
tidy_pkgs <- tidyverse::tidyverse_packages(include_self = FALSE)
print(tidy_pkgs)
```
We can Install all these packages in the tidyverse by running `install.packages("tidyverse")` .
Among these `r length(tidy_pkgs)` packages, **ggplot2**, **dplyr**, **tidyr**,
**readr**, **purrr**, **tibble**, **stringr**, and **forcats** are the so-called *core packages*, which will be loaded all together if we run `library(tidyverse)`. See this: https://www.tidyverse.org/packages/

These articles may be of our interest:

- https://support.posit.co/hc/en-us/articles/201057987-Quick-list-of-useful-R-packages
- https://www.datacamp.com/tutorial/top-ten-most-important-packages-in-r-for-data-science



## Writing R Packages

