# Lesson 4 Data Structures {#L4}

**Acknowledgments:** This chapter is written after I read Chapter 3 of Hadley Wickham's book, *Advanced R* 2nd Ed, https://adv-r.hadley.nz/vectors-chap.html


In R, the fundamentally important data structure is **vector,** which can be roughly defined as *a tuple of elements.* R vectors include **atomic vectors** and **lists**^[According to Hadley Wickham, *Advanced R* 2nd Ed., "`NULL` is closely related to vectors and often serves the role of a generic zero length vector."]. Other R data structures are built upon vectors by adding more *attributes.* Talking about attributes, I want to quote words from Hadley Wickham (*Advanced R* 2nd Ed, https://adv-r.hadley.nz/vectors-chap.html):

>
The most important attributes are names, dimensions and class.
>

>
Two attributes are particularly important. The dimension attribute turns vectors into matrices and arrays and the class attribute powers the S3 object system.
>


## Atomic Vectors

The most useful atomic vectors are **logical**, **integer**, **double** and **character** types.

**Examples:**
```{r}
(a_logic_vec <- c(TRUE, FALSE, TRUE))
(an_integer_vec <- 1L:6L)
(a_double_vec <- c(1, 2, 3, 4, 5, 6))
(identical(an_integer_vec, a_double_vec))
(a_character_vec <- c('a', 'b', 'c', '123'))
```

We can use R function `typeof()` to find out the type of an atomic vector.
```{r}
c(typeof(a_logic_vec), 
  typeof(an_integer_vec), 
  typeof(a_double_vec),
  typeof(a_character_vec))
```
or
```{r}
vapply(list(a_logic_vec, 
            an_integer_vec, 
            a_double_vec, 
            a_character_vec), typeof, character(1L))
```

**Important:**

1. When different types of atomic-vector variables are on "operation", R automatically changes variables' types. The rule of change is: from *low* to just *high* enough. As for "low" and "high", the order is
$$
\hbox{logical}<\hbox{integer}<\hbox{double}<\hbox{character}
$$
Note that when converting from logical to integer, `TRUE` and `FALSE` become `1L` and `0L`, respectively.
1. R function `length()` gives *length*---how many elements an atomic vector has.
1. When two atomic-vector variables of different lengths are on "operation", R will firstly recycle the shorter variable.
1. We use `vec[index]` or other ways^[For details see Hadley Wickham, https://adv-r.hadley.nz/subsetting.html]  to get values from an atomic vector.

**Examples:**
```{r}
(a_logic_vec + an_integer_vec)
print(c(an_integer_vec, a_double_vec))
print(c(a_logic_vec, a_double_vec, a_character_vec))
print(a_character_vec[c(1, 4)])
```

```{r}
a <- c(1, 1e3, 1e5)
b <- c(1L, 1e3L, 1e5L)
(a == b) # type changes when on operation
(identical(a, b))
(as.character(a)) # interesting!
(as.character(b))
(as.character(as.integer(a))) # this might be useful in practice.
```


## Lists

A list and an atomic vector share the common property of "a tuple of elements." A list differs an atomic vector in that each element (item) in a list can be **more complex stuff** but not limited to a logical/integer/double/character value. We use 
`a_list[index]` to subset a list, and we use 
`a_list[[index]]` or `a_list[[item name]]` to have **one item** from a list.
Let's have an example
```{r}
a_list <- list(item_1 = 1L:3L,
               item_2 = "I like R",
               item_3 = c(TRUE, FALSE),
               item_4 = c(1.8, 2.6, 3.3, 8.9, 10.0))
(is.list(a_list))
print(names(a_list))
print(length(a_list))
print(lengths(a_list))
print(a_list[c(1, 2)])
print(a_list[2])
print(a_list[[2]])
print(a_list[["item_2"]])
```

## Matrices

A matrix is a vector with the 'dim' attribute.
```{r}
a_matrix <- 
  structure(1:9,
            dim = c(3, 3))
(is.matrix(a_matrix))
(attributes(a_matrix))
print(a_matrix)

b_matrix <- matrix(1:9, ncol = 3)
(is.matrix(b_matrix))
(attributes(b_matrix))
print(b_matrix)

(identical(a_matrix, b_matrix))
```


## Data Frames

A data frame is a rectangular table with columns and rows. Technically speaking,
a data frame is a **list of atomic vectors**---they have the same length---plus three attributes: a) 'names' (for columns); b) 'row.names'; and 'class =  “data.frame”' 
**Data frames share the properties of both matrices and lists.**

```{r}
my_df_1 <- 
  structure(list(1:5,
                 letters[1:5],
                 c(TRUE, FALSE, TRUE, FALSE, TRUE)),
            names = c('a', 'b', 'c'),
            row.names = 1:5,
            class = "data.frame")
is.data.frame(my_df_1)
attributes(my_df_1)

my_df_2 <- data.frame(a = 1:5, 
                      b = letters[1:5],
                      c = c(TRUE, FALSE, TRUE, FALSE, TRUE),
                      row.names = 1:5,
                      stringsAsFactors = FALSE)
identical(my_df_1, my_df_2)
```

## Factors and Dates

Again, I quote words from Hadley Wickham (*Advanced R* 2nd Ed, https://adv-r.hadley.nz/vectors-chap.html):

About factors

>
Factors are built on top of an integer vector with two attributes: a class, “factor”, which makes it behave differently from regular integer vectors, and levels, which defines the set of allowed values.
>

>
It’s usually best to explicitly convert factors to character vectors if you need string-like behaviour.
>

About Dates

>
Date vectors are built on top of double vectors. They have class “Date” and no other attributes.
>

